ITS_LIVE: A Cloud-Native Approach to Monitoring Glaciers from Space

### 2023-11-28 paper.qmd
updating text

ITS_LIVE users, like [most Earth Scientists @Fouilloux2018-jw;@Marshall2022-je;@Lopez2021-el], primarily access the raw data in two ways: (1) spatial exploration and (2) temporal exploration. Spatial exploration is supported through library access to the millions of NetCDF granules that can be identified programmatically using an OpenAPI search or interactively using the ITS_LIVE webtool. Temporal exploration is accommodated by a restructuring of the NetCDF granules into Zarr datacubes that are composed of 10 x 10 x 20,000 chunked data, where 10 x 10 is in the spatial domain and 20,000 is in the temporal domain. This chunking scheme minimizes the amount of data that needs to be uncompressed for a time series at a single location, providing near-zero latency for temporal exploration, which is critical for process understanding and change detection.

@@

ITS_LIVE provides users with intuitive no-code options to explore the rich dataset through a Voila-based dashboard and a serverless application [Fig. @fig-3] In the following pages we present some of the data visualizations we can obtain from using ITS_LIVE data without costly pre-processing steps using open-source GIS tools like QGIS[8] and the Pangeo stack in Python ([Xarray @Hoyer2017-su], [Zarr @Miles2023-yj] , [HoloViews @Rudiger2020-yr]). A companion GitHub repository contains all the notebooks and QGIs projects used to generate the figures in this article. The focus areas include Malaspina Glacier in Alaska (@fig-2;@fig-5), Pine Island Glacier in Antarctica and Antarctic ice shelves (@fig-4)

@@

[See @Miles2023-yj;@Fouilloux2018-jw;@Scambos1992-jz]
:::

@@

What looks like a frequency spectrogram on the right in  this figure is a velocity profile along sampled points in Malaspina glacier in Alaska (above) , with the first point at the lower end of the ice (0 km at the left edge). In this plot the ice is flowing from the right to the left (A to B). One can see the seasonal cycle in speed (slowest in Fall) in the upper glacier, and the surges of the Malaspina in indicated years, including 2020 and 2021, when the entire glacier along this centerline reached speeds > 2500 m/yr, or 6.8 m/day.

>>>>

ITS_LIVE users, like [most Earth Scientists @abernathey2021cloud;@Marshall2022-je], primarily access the raw data in two ways: (1) spatial exploration and (2) temporal exploration. Spatial exploration is supported through library access to the millions of NetCDF granules that can be identified programmatically using an OpenAPI search or interactively using the ITS_LIVE webtool. Temporal exploration is accommodated by a restructuring of the NetCDF granules into Zarr datacubes that are composed of 10 x 10 x 20,000 chunked data, where 10 x 10 is in the spatial domain and 20,000 is in the temporal domain. This chunking scheme minimizes the amount of data that needs to be uncompressed for a time series at a single location, providing near-zero latency for temporal exploration, which is critical for process understanding and change detection.

@@

ITS_LIVE provides users with intuitive no-code options to explore the rich dataset through a Voila-based dashboard and a serverless application [Fig. @fig-3] In the following pages we present some of the data visualizations we can obtain from using ITS_LIVE data without costly pre-processing steps using open-source GIS tools like QGIS[8] and the Pangeo stack in Python ([Xarray @Hoyer2017-su], [Zarr @Miles2023-yj] , etc). A companion GitHub repository contains all the notebooks and QGIs projects used to generate the figures in this article. The focus areas include Malaspina Glacier in Alaska (@fig-2;@fig-5), Pine Island Glacier in Antarctica and Antarctic ice shelves (@fig-4)

@@

[See @Miles2023-yj;@Scambos1992-jz]
:::

@@

What looks like a frequency spectrogram on the right in  this figure is a velocity profile along sampled points in Malaspina glacier in Alaska (above) , with the first point at the lower end of the ice (0 km at the left edge). In this plot the ice is flowing from left to right (A to B). One can see the seasonal cycle in speed (slowest in Fall) in the upper glacier, and the surges of the Malaspina in indicated years, including 2020 and 2021, when the entire glacier along this centerline reached speeds > 2500 m/yr, or 6.8 m/day.
